---
title: "Data Science for Public Policy"
subtitle: "Assignment 04: Data Visualization, Quarto, and Git"
author: "Juliette Taylor - jmt333, Thomas Tiberghien Alvarez - tt791"
format: 
  html:
    code-line-numbers: true
execute: 
  warning: false
urlcolor: blue
---

\vspace{0.1in}

```{r}
#| output: false
library(tidyverse)
library(knitr)
library(lubridate)
library(purrr)
```


The code below is used to download the data zip files into csvs that are then ingested by R. Make sure that the period input is equivalent to the ones listed in https://s3.amazonaws.com/capitalbikeshare-data/index.html
```{r}
cabi_csv <- function(period, zip_dir = "data/", csv_dir = "csvs/") {
        url <- str_glue("https://s3.amazonaws.com/capitalbikeshare-data/{period}-capitalbikeshare-tripdata.zip")
        download.file(url, destfile = str_glue(paste0(zip_dir,"{period}-cabi-tripdata.zip")))
        zip_files <- paste0(zip_dir, list.files(zip_dir))
        for (x in zip_files) {
                unzip(x, exdir = csv_dir)
        }
        csv_files <- paste0(csv_dir, list.files(csv_dir))
        output <- map(csv_files, read_csv)
        file.remove(zip_files)
        file.remove(csv_files)
        return(bind_rows(output))
}
```

## Plot 1: Yearly changes

The goal of this plot is to simply see how ridership has changed throughout the last 10 years. The output graph is relatively simple, but in order to obtain the data, several data sources had to be wrangled together. For simplicity, only data from May will be used. 

